Expanding the Power of Lens Synthesis \\ (talk proposal)

### 2019-02-11 papers/BXProposal/main.tex
Edits to BX proposal

\title{Expanding the Power of Lens Synthesis With Annotations}

% I've never done a talk before, should this be everybody or just me
\author{
Anders Miltner \\ Princeton University\\
Princeton, NJ 08540 \\ amiltner@cs.princeton.edu
\and
Solomon Maina \\ University of Pennsylvania\\
Philadelphia, PA 19104 \\ smaina@seas.upenn.edu
\and
Kathleen Fisher \\ Tufts University\\
Medford, MA 02155 \\ kfisher@eecs.tufts.edu
\and
Benjamin C. Pierce \\ University of Pennsylvania\\
Philadelphia, PA 19104 \\ bcpierce@cis.upenn.edu
\and
David Walker \\ Princeton University\\
Princeton, NJ 08540 \\ dpw@cs.princeton.edu
\and
Steve Zdancewic \\ University of Pennsylvania\\
Philadelphia, PA 19104 \\ stevez@cis.upenn.edu
}

@@

\begin{abstract}
  Recent work has shown how to synthesize a class of string bijections from
  regular expression specifications. While these bijective string
  transformations can be useful for synchronizing data formats, it may still be
  useful to keep interesting formats synchronized when they are not in bijective
  correspondence. In this talk, we aim to show how we can expand the class of
  synthesized transformations beyond bijections, as long as users are willing to
  provide annotations that help identify a ``bijective core.''
\end{abstract}
\vskip 32pt

@@

Bidirectional languages are often difficult to program in. The can use
unfamiliar paradigms, and their type systems can sometimes be complex, making
fixing type errors oftentimes nontrivial. This is evident in the language
Boomerang~\cite{boomerang}, a DSL for writing lenses~\cite{Focal2005-long2}
on strings. In Boomerang, users can write bidirectional programs between data
formats: programs are lenses and types are regular expressions representing the
formats the lenses operate on. Instead of requiring users to manually write
bidirectional transformations, we find it preferable to synthesize bidirectional
transformations from their types.

Optician is a Boomerang extension that synthesizes a subset of Boomerang lenses,
bijective lenses. When provided with a pair of regular expressions, Optician
will synthesize one of Boomerang's bijective lenses that converts between the
provided regular expressions.

>>>>

\title{Expanding the Power of Lens Synthesis \\ (talk proposal)}

% I've never done a talk before, should this be everybody or just me
\author{
Anders Miltner \\ Princeton University \\ \ 
% Princeton, NJ 08540 \\ amiltner@cs.princeton.edu
\and
Solomon Maina \\ University of Pennsylvania
% Philadelphia, PA 19104 \\ smaina@seas.upenn.edu
\and
Kathleen Fisher \\ Tufts University
% Medford, MA 02155 \\ kfisher@eecs.tufts.edu
\and
Benjamin C. Pierce \\ University of Pennsylvania
% Philadelphia, PA 19104 \\ bcpierce@cis.upenn.edu
\and
David Walker \\ Princeton University
% Princeton, NJ 08540 \\ dpw@cs.princeton.edu
\and
Steve Zdancewic \\ University of Pennsylvania
% Philadelphia, PA 19104 \\ stevez@cis.upenn.edu
}

@@

\begin{abstract}
Recent work has shown how to synthesize a class of string bijections from
regular expression specifications. These bijective string transformations
can be useful for synchronizing simple data formats, but many interesting
formats require non-bijective lenses. In this talk, we show how to expand
the class of synthesized transformations beyond bijections to a large subset
of full symmetric lenses (including all the standard asymmetric lenses)
using lightweight user annotations that help identify a ``bijective core.''
\end{abstract}

@@

Bidirectional languages can be difficult to program in, proposing unfamiliar
programming paradigms, novel syntax, and subtle typing constraints.
Fortunately, their restricted expressive power, together with the
fine-grained type systems needed to guarantee that well-types programs
satisfy strong round-tripping properites, make them a perfect target for
program synthesis.

The {\em Optician} tool is an extension to Boomerang~\cite{boomerang}---a
DSL for writing lenses~\cite{Focal2005-long2} on strings---that synthesizes
{\em bijective} Boomerang lenses when provided with a pair of types (regular
expressions) and a small number of examples (often zero).
